Import data from the Toronto Paramedic Service via Open Data Toronto

### 2023-02-03 outputs/paper/paper.qmd
General updates and bug fixes
* Moved data cleaning to a separate script file
* Renamed emergent care data set to a more meaningful title: performance data
* Updated readme.md

```{r}
#| messages: false
#| echo: false
#| warning: false
#### Workspace set-up ####
# Define required libraries 
library(lubridate)
library(tidyverse)
library(dplyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(readr)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#### Fetch and clean relevant data sets ####
# Fetch the .csv file created from Open Data Toronto's "Land Ambulance Response Time Standard" package: https://open.toronto.ca/dataset/land-ambulance-response-time-standard
raw_response_data <- readr::read_csv(here::here("inputs/data/raw_response_data.csv"), show_col_types = FALSE) |>
  clean_names()

# Fetch Open Data Toronto's "Pre-Hospital Emergency Care Performance Metrics" package:  https://open.toronto.ca/dataset/pre-hospital-emergency-care-performance-metrics
raw_emergent_care_data <- readr::read_csv(here::here("inputs/data/raw_emergent_care_data.csv"), show_col_types = FALSE) |>
  clean_names()

# Fetch Open Data Toronto's "Paramedic Services Incident Data" package: https://open.toronto.ca/dataset/paramedic-services-incident-data
raw_incident_data <- readr::read_csv(here::here("inputs/data/raw_incident_data.csv"), show_col_types = FALSE) |>
  clean_names()
```

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Number of Emergency Medical Responses in Toronto (by year since 2014)"
#| label: fig-1
raw_emergent_care_data |>
  # Define plot 
  ggplot(aes(
    x = year, 
    y = actual
  )) +

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Emergency Medical Incidents and Dispatched Ambulances (in 2021)"
#| label: fig-2
clean_incident_data <- raw_incident_data |>
  # Group data by incident type
  group_by(incident_type) |>

@@

# Rename "-" in data to more verbose label
clean_incident_data[1, "Incident Type"] = "Other Incident"

# Render table
clean_incident_data |>
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>
  add_header_above(
    c(" ", "Year Totals" = 2, "Dispatch Averages" = 2), 
    align = "c",
    bold=TRUE
  ) |>
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE) |>
  row_spec(0,bold=TRUE) |>
  column_spec(1, width = "13em") 
```

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Mandated Response Times for the Toronto Paramedic Services (as of 2021)"
#| label: fig-3
raw_response_data |>
  # Select type of response and estimated response plan 
  select(patient_category, plan_in_minutes) |>

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Percent of Calls Meeting Mandated Response Times (since 2013)"
#| label: fig-4
raw_response_data |>
  # Select desired variables
  select(patient_category, performance_in_percentage, plan_in_percentage, response_time_standard_year) |>

  # Remove rows with no data
  filter(performance_in_percentage != "NA") |>

  # Plot graphs, define x axis
  ggplot(aes(x=response_time_standard_year)) +

@@

  # Merge graphs into one figure
  facet_wrap(~patient_category)

>>>>

```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Workspace set-up ####
# Define required libraries 
library(lubridate)
library(tidyverse)
library(dplyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(readr)
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#### Fetch and clean relevant data sets ####
# Fetch the the clean response data
clean_response_data <- readr::read_csv(here::here("inputs/data/clean_response_data.csv"), show_col_types = FALSE)

# Fetch clean performance data
clean_performance_data <- readr::read_csv(here::here("inputs/data/clean_performance_data.csv"), show_col_types = FALSE)

# Fetch the clean incident response
clean_incident_data <- readr::read_csv(here::here("inputs/data/clean_incident_data.csv"), show_col_types = FALSE)
```

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Number of Emergency Medical Responses in Toronto (by year since 2014)"
#| label: fig-1
clean_performance_data |>
  # Define plot 
  ggplot(aes(
    x = year, 
    y = ems_responses
  )) +

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Emergency Medical Incidents and Dispatched Ambulances (in 2021)"
#| label: fig-2
summarized_incident_data <- clean_incident_data |>
  # Group data by incident type
  group_by(incident_type) |>

@@

# Rename "-" in data to a more meaningful label
summarized_incident_data[1, "Incident Type"] = "Other Incident"

# Order rows alphabetically by incident type
summarized_incident_data <- summarized_incident_data[order(summarized_incident_data$`Incident Type`),]

# Render table
summarized_incident_data |>
  # Table configuration 
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>

  # Add headers for table data
  add_header_above(
    c(" ", "Year Totals" = 2, "Dispatch Averages" = 2), 
    align = "c",
    bold = TRUE
  ) |>

  # Style the table 
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE) |>

  # Make the table headers bold
  row_spec(0,bold=TRUE) |>

  # Increase width of first column
  column_spec(1, width = "13em") 
```

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Mandated Response Times for the Toronto Paramedic Services (as of 2021)"
#| label: fig-3
clean_response_data |>
  # Select type of response and estimated response plan 
  select(patient_category, plan_in_minutes) |>

@@

```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Percent of Calls Meeting Mandated Response Times (since 2013)"
#| label: fig-4
clean_response_data |>
  # Plot graphs, define x axis
  ggplot(aes(x=response_time_standard_year)) +

@@

  guides(shape=TRUE) +

  # Merge graphs into one figure
  facet_wrap(~patient_category) 
```

### 2023-02-03 scripts/01-data_import.qmd
General updates and bug fixes
* Moved data cleaning to a separate script file
* Renamed emergent care data set to a more meaningful title: performance data
* Updated readme.md

```{r}
# Fetch Open Data Toronto's "Land Ambulance Response Time Standard" package: https://open.toronto.ca/dataset/land-ambulance-response-time-standard 
raw_response_data <-
  list_package_resources("8b068db1-95b8-4ea9-b942-8a7ce3a5fbd6") |>
  filter(name == "Response Time (2013 - 2021)") |>
  get_resource()

@@

```{r}
# Fetch Open Data Toronto's "Pre-Hospital Emergency Care Performance Metrics" package:  https://open.toronto.ca/dataset/pre-hospital-emergency-care-performance-metrics
raw_emergent_care_data <-
  list_package_resources("174ace54-bcc7-4885-85c1-76094ab4a0ca") |>
  filter(name == "Pre-hospital emergency care performance metrics 2014-2021") |>
  get_resource() |>
  filter(`Pre-Hospital Emergency Care Type` == "Emergency Request for Service")

# Write data to a .csv file
write_csv(
  x = raw_emergent_care_data,
  file = "../inputs/data/raw_emergent_care_data.csv"
)
```

```{r}
# Fetch Open Data Toronto's "Paramedic Services Incident Data" package: https://open.toronto.ca/dataset/paramedic-services-incident-data
raw_incident_data <-
  list_package_resources("c21f3bd1-e016-4469-abf5-c58bb8e8b5ce") |>
  filter(name == "paramedic-services-incident-data-2017-2021") |>
  get_resource()

# Write data to a .csv file
write_csv(
  x = raw_incident_data$`2021`,
  file = "../inputs/data/raw_incident_data.csv",
)
```

>>>>

```{r}
#| warning: false
# Fetch Open Data Toronto's "Land Ambulance Response Time Standard" package: https://open.toronto.ca/dataset/land-ambulance-response-time-standard 
raw_response_data <-
  list_package_resources("8b068db1-95b8-4ea9-b942-8a7ce3a5fbd6") |>
  filter(name == "Response Time (2013 - 2021)") |>
  get_resource()

@@

```{r}
#| warning: false
# Fetch "Pre-Hospital Emergency Care Performance Metrics" from Open Data Toronto
# Link: https://open.toronto.ca/dataset/pre-hospital-emergency-care-performance-metrics
raw_performance_data <-
  list_package_resources("174ace54-bcc7-4885-85c1-76094ab4a0ca") |>
  filter(name == "Pre-hospital emergency care performance metrics 2014-2021") |>
  get_resource() |>
  filter(`Pre-Hospital Emergency Care Type` == "Emergency Request for Service")

# Write data to a .csv file
write_csv(
  x = raw_performance_data,
  file = "../inputs/data/raw_performance_data.csv"
)
```

```{r}
#| warning: false
# Fetch Paramedic Services Incident Data from Open Data Toronto
# Link: https://open.toronto.ca/dataset/paramedic-services-incident-data
raw_incident_data <-
  list_package_resources("c21f3bd1-e016-4469-abf5-c58bb8e8b5ce") |>
  filter(name == "paramedic-services-incident-data-2017-2021") |>
  get_resource()

# Write data to a .csv file
write_csv(
  x = raw_incident_data$`2021`,
  file = "../inputs/data/raw_incident_data.csv",
)

# Note: creating this .csv may take some time as it's a large data set
```
